Automated Metadata Generation for Fish Specimen Image Collections

### 2021-03-24 paper/paper.tex
Updates from Overleaf
First draft Pixel Analysis section

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Pepper, Breen and Greenberg}

@@

\subsection{Metadata genberation?}

@@

\section{Methods}
Text
\section{Results}
Text
\section{Discussion}
Text \textit{texty text}%~\cite{Lamport:LaTeX}
\section{Conclusion}
Text \textbf{text}

>>>>

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Pepper, Greenberg and Breen}

@@

\subsection{Metadata generation?}

@@

\section{Methods}

\subsection{Detectron}
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they are almost never perfect. They almost always either include additional background as part of the fish, or clip part(s) of the fish outside the bounding box. To solve this shortcoming, we utilize pixel analysis methods commonly found in image informatics.

The first calculation in the pixel analysis process is determining the cutoff intensity between what constitutes the foreground (i.e.\ the fish) and background of the image. Initially, we base this off the bounding box and mask generated by \verb|detectron|. Specimen images are read in as gray scale, and pixels in the image are treated as unsigned integers between 0 and 255. Otsu thresholding is used to get an initial cutoff value between foreground and background. \todo{Put in another sentence or two about Otsu} While this value occasionally generates an accurate mask as is, most of the time the contrast between foreground and background is not great enough and much of the lighter parts of the fish (such as its tail fin) are marked as background.

To overcome this, the threshold value needs either scaled up or down depending on if the background is lighter or darker than the fish. For our current dataset, the background is always lighter (i.e.\ closer to 255), so the threshold value needs scaled up to include more of the image as foreground. Precisely how much to scale this value by is difficult to determine. For optimal results it is dependent on the level of contrast between the background and foreground, which can be affected by both how washed out the image is and how light or dark the pigmentation of the fish is. To take this into account, we measure the mean intensity of the foreground and background, then use these intensity values and the difference between them to scale the threshold value. Even with these values in hand it is not entirely clear how to scale the threshold, but we have found\ldots{}\todo{This is still somewhat indeterminate, need to pick a definitive technique, I've tried a few}

While thresholding has the potential to generate better masks than a neural network (when provided an initial approximate bounding box), it also introduces considerable noise. Single or small groups of errant pixels can be marked as foreground depending on the consistency of the background, and interior pixels of the fish (especially around the fins) can be marked as background. To be useful for generating an accurate bounding box and for subsequent computational analysis, the mask must consist of one single ``blob'' over the fish the contains no holes, and no other pixels disconnected from this blob can be marked as foreground.

To accomplish this, we apply an iterative process of flood filling from all the foreground pixels in the image until a blob is generated that is large enough to constitute the fish. This is another metaparameter, but greater than 10\% of the current bounding box has masked the specimen in all observed cases. Once the fish's blob is found, noise then needs removed. This is done by flood filling from each of the corner of the bounding box where the specimen is not present (all four corners in the overwhelming majority of cases), then taking the inverse of the result. The fish mask is excluded from these corner flood fills, so this process removes all noise from both the background and foreground of the image leaving only a close mask over the fish itself.

With an accurate mask generated, it is then necessary to check whether the bounding box needs expanded or shrunk along any edges. Expansion is done first, buy checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole the process, masking and expansion process is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the mask is the shrunk. This process is more efficient as all that is required is to shrink each edge by one up until the point they contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.
\subsection{Metadata Generation}
\section{Results}
Text
\section{Discussion}
Text \textit{texty text}%~\cite{Lamport:LaTeX}
\section{Conclusion}
Text \textbf{text}
